---
title: "Diagnose"
author: "Catherine"
date: "12/9/2018"
output: 
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
library(tidyverse)
library(patchwork)
library(HH)
```

# fit a model for diagnose

# DATA Preprocessing

Import original dataset:
```{r import_data_p2, include = FALSE, message = FALSE}
data = read_csv(file = './Cancer_Registry.csv') %>% 
  janitor::clean_names()
```

build 'region' variable from 'geography'. This is gonna be the only categorical variable in the dataset.
```{r include = FALSE}
data_pre = data %>% 
separate(., geography, into = c('county', 'state'), sep = ', ') %>% 
  mutate(., region = replace(state, state == 'Connecticut', 'Northeast'),
            region = replace(region, region == 'Maine', 'Northeast'),
            region = replace(region, region == 'Massachusetts', 'Northeast'),
            region = replace(region, region == 'New Hampshire', 'Northeast'),
            region = replace(region, region == 'Rhode Island', 'Northeast'),
            region = replace(region, region == 'Vermont', 'Northeast'),
            region = replace(region, region == 'New Jersey', 'Northeast'),
            region = replace(region, region == 'New York', 'Northeast'),
            region = replace(region, region == 'Pennsylvania', 'Northeast'),
            region = replace(region, region == 'Indiana', 'Midwest'),
            region = replace(region, region == 'Michigan', 'Midwest'),
            region = replace(region, region == 'Illinois', 'Midwest'),
            region = replace(region, region == 'Ohio', 'Midwest'),
            region = replace(region, region == 'Wisconsin', 'Midwest'),
            region = replace(region, region == 'Iowa', 'Midwest'),
            region = replace(region, region == 'Kansas', 'Midwest'),
            region = replace(region, region == 'Minnesota', 'Midwest'),
            region = replace(region, region == 'Missouri', 'Midwest'),
            region = replace(region, region == 'Nebraska', 'Midwest'),
            region = replace(region, region == 'North Dakota', 'Midwest'),
            region = replace(region, region == 'South Dakota', 'Midwest'),
            region = replace(region, region == 'Delaware', 'South'),
            region = replace(region, region == 'Florida', 'South'),
            region = replace(region, region == 'Georgia', 'South'),
            region = replace(region, region == 'Maryland', 'South'),
            region = replace(region, region == 'North Carolina', 'South'),
            region = replace(region, region == 'South Carolina', 'South'),
            region = replace(region, region == 'Virginia', 'South'),
            region = replace(region, region == 'District of Columbia', 'South'),
            region = replace(region, region == 'West Virginia', 'South'),
            region = replace(region, region == 'Alabama', 'South'),
            region = replace(region, region == 'Kentucky', 'South'),
            region = replace(region, region == 'Mississippi', 'South'),
            region = replace(region, region == 'Tennessee', 'South'),
            region = replace(region, region == 'Arkansas', 'South'),
            region = replace(region, region == 'Louisiana', 'South'),
            region = replace(region, region == 'Oklahoma', 'South'),
            region = replace(region, region == 'Texas', 'South'),
            region = replace(region, region == 'Arizona', 'West'),
            region = replace(region, region == 'Colorado', 'West'),
            region = replace(region, region == 'Idaho', 'West'),
            region = replace(region, region == 'Montana', 'West'),
            region = replace(region, region == 'Nevada', 'West'),
            region = replace(region, region == 'New Mexico', 'West'),
            region = replace(region, region == 'Utah', 'West'),
            region = replace(region, region == 'Wyoming', 'West'),
            region = replace(region, region == 'Alaska', 'West'),
            region = replace(region, region == 'California', 'West'),
            region = replace(region, region == 'Hawaii', 'West'),
            region = replace(region, region == 'Oregon', 'West'),
            region = replace(region, region == 'Washington', 'West'))
```

```{r include = FALSE}
data_pre = 
  mutate(data_pre, pct_some_col18_24 = 100 - (pct_no_hs18_24 + pct_hs18_24 + pct_bach_deg18_24))
```

```{r include = FALSE}
data_vrb_pre_selec = data_pre %>%
  dplyr::select(., -'avg_deaths_per_year', -'binned_inc', -'pct_married_households', -'pct_employed16_over', -'pct_private_coverage_alone', -'county', -'state')
```

## The model we got from model selection is 

`target_death_rate ~ incidence_rate + poverty_percent + median_age_female + region + pct_hs18_24 + pct_hs25_over + pct_unemployed16_over + pct_public_coverage, data = data_vrb_pre_selec`

```{r}
final_fit = lm(target_death_rate ~ incidence_rate + poverty_percent + median_age_female + region + pct_hs18_24 + pct_hs25_over + pct_unemployed16_over + pct_public_coverage, data = data_vrb_pre_selec) 
```

# Then we take a look at the histogram of varibles

```{r histogram, include = FALSE}
p = data_vrb_pre_selec %>% 
  ggplot(aes(target_death_rate)) + geom_histogram()

p1 = data_vrb_pre_selec %>% 
  ggplot(aes(poverty_percent)) + geom_histogram()

p2 = data_vrb_pre_selec %>% 
  ggplot(aes(median_age_female)) + geom_histogram()

p3 = data_vrb_pre_selec %>% 
  ggplot(aes(pct_hs18_24)) + geom_histogram()

p4 = data_vrb_pre_selec %>% 
  ggplot(aes(pct_hs25_over)) + geom_histogram()

p5 = data_vrb_pre_selec %>% 
  ggplot(aes(pct_unemployed16_over)) + geom_histogram()

p6 = data_vrb_pre_selec %>% 
  ggplot(aes(pct_public_coverage)) + geom_histogram()

p7 = data_vrb_pre_selec %>% 
  ggplot(aes(incidence_rate)) + geom_histogram()

p

(p1 + p2 + p3) / (p4 + p5 + p6)

t1 = data_vrb_pre_selec %>% 
  ggplot(aes(log(poverty_percent))) + geom_histogram()

t1 + p1
```

```{r transformation}
log_final_fit = lm(target_death_rate ~ incidence_rate + log(poverty_percent) + median_age_female + region + pct_hs18_24 + pct_hs25_over + pct_unemployed16_over + pct_public_coverage, data = data_vrb_pre_selec) 
```

__Comment__

We found that `poverty_percent` is right-skewed so we decided to do transformation for this variable. 

We got our model after transformation: `target_death_rate ~ incidence_rate + log(poverty_percent) + median_age_female + region + pct_hs18_24 + pct_hs25_over + pct_unemployed16_over + pct_public_coverage`

# plot the regression model

```{r}
par(mfrow = c(2,2))
plot(log_final_fit)
```

# Detect outliers in Y using ‘studentized residuals’

```{r}
sr = rstandard(log_final_fit) %>% as_tibble() %>% mutate(n = 1:3047)
outlier_y = sr %>% filter(abs(value) > 2.5) 
names(outlier_y) = c("over_2.5", "observation_n") 
knitr::kable(outlier_y, caption = "Outliers of y")
```

__Comment__

There are __70__ ouliers in Y being detected. 

# Detect outliners in X using Leverage values

```{r}
fit_hat = hatvalues(log_final_fit) %>% as_tibble() %>% mutate(n = 1:3047)
outlier_x_moderate = fit_hat %>% filter(abs(value) > 0.2) 
outlier_x_high = fit_hat %>% filter(abs(value) > 0.5)
names(outlier_x_moderate) = c("over_0.2", "observation_n") 
names(outlier_x_high) = c("over_0.5", "observation_n")
knitr::kable(outlier_x_moderate, caption = "Outliers of X (> 0.2)")
knitr::kable(outlier_x_high, caption = "Outliers of X (> 2p/n)")
```

__Comment__

By taking look at the $h_{ii}$ values, we detect __no oulier in X__ with both cutoff $h_{ii} > 0.2$ and cutoff $h_{ii} > 0.5$.

# Influencial Observation

Not all outliers are influential. Therefore, we need to test the influence of the outliers. 
Using DFFITS test the difference of fitted value with/without an observation and Cook's Distance to find concerned values.

```{r}
tb = influence.measures(log_final_fit)[["infmat"]] %>% as_tibble() %>% 
  mutate(n = 1:3047) %>% 
  select(dffit, cook.d, n) %>% 
  filter(abs(dffit)>1|abs(cook.d)>0.5)
knitr::kable(tb, caption = "Influential observation")
```

__Comment__

Consider `DFFITS` and `Cook's Distance`, we found out an influencitial outlier `282`. The difference is large between fitted value with/without `282` observation. Next, we can take a look at the change of fitted value with/without `282`.

```{r}
without282 = data_vrb_pre_selec[-282,]
fit_with282 = lm(target_death_rate ~ incidence_rate + log(poverty_percent) + median_age_female + region + pct_hs18_24 + pct_hs25_over + pct_unemployed16_over + pct_public_coverage, data = data_vrb_pre_selec)
fit_without282 = lm(target_death_rate ~ incidence_rate + log(poverty_percent) + median_age_female + region + pct_hs18_24 + pct_hs25_over + pct_unemployed16_over + pct_public_coverage, data = without282)

sum1 = summary(fit_with282)$coef 
sum2 = summary(fit_without282)$coef

knitr::kable(sum1, caption = "Model with observation 282")
knitr::kable(sum2, caption = "Model without observation 282")



coef2 = (sum1[2]-sum2[2])/sum1[2]*100
coef3  = (sum1[3]-sum2[3])/sum1[3]*100
coef4 = (sum1[4]-sum2[4])/sum1[4]*100
coef5 = (sum1[5]-sum2[5])/sum1[5]*100
coef6 = (sum1[6]-sum2[6])/sum1[6]*100
coef7 = (sum1[7]-sum2[7])/sum1[7]*100
coef8 = (sum1[8]-sum2[8])/sum1[8]*100
coef9 = (sum1[9]-sum2[9])/sum1[9]*100
coef10 = (sum1[10]-sum2[10])/sum1[10]*100 
coef11 = (sum1[11]-sum2[11])/sum1[11]*100

rbind(incidence_rate = coef2, log_poverty_percent = coef3, median_age_female = coef4, regionNortheast = coef5, regionSouth = coef6, regionWest = coef7, pct_hs18_24 = coef8, pct_hs25_over = coef9, pct_unemployed16_over = coef10, pct_public_coverage = coef11) 

```


__Comment__

After calculating the coefficient changes for each variables, we found that the changes are not significant. We decided to keep the observation `282` in our model. 


```{r fig.height=6}
fit_with282 = lm(target_death_rate ~ incidence_rate + log(poverty_percent) + median_age_female + region + pct_hs18_24 + pct_hs25_over + pct_unemployed16_over + pct_public_coverage, data = data_vrb_pre_selec)

par(mfrow = c(2,2))
plot(fit_with282)
```

__Comment__

In the _Residuals vs Fitted Plot_ and _Scale-Location Plot_, residual values are roughly bounce around 0. 

In the _Quantile–Quantile Plot_, it is almost a straight line and no heavy tails. Small departures from normality are not concerning to our model. 

In the _Residuals vs Leverage Plot_, there is no outlying values at the upper right or lower right corner.


# Multicolinearity

```{r}
cor_fit = data_vrb_pre_selec %>% 
  dplyr::select(target_death_rate, incidence_rate, poverty_percent, median_age_female, pct_hs18_24, pct_hs25_over, pct_unemployed16_over, pct_public_coverage)
round(cor(cor_fit),3) %>% knitr::kable()

vif(fit_without282)
```

__Comments__

By checking Correlation Matrix and vif value of each varibles, we found no varibles 

