---
title: "Validation"
author: "Baoyi Shi"
output: html_document
---
```{r import_packages, echo = FALSE, message = FALSE}
library(tidyverse)
library('leaps')
library(faraway)
library(broom)
library('leaps')
library(caret)
library(dplyr)
library(readxl)
```


=======================================================================================================

# DATA Preprocessing

Import original dataset:
```{r import_data_p2, echo = FALSE, message = FALSE}
data = read_csv(file='./Cancer_Registry.csv') %>% 
  janitor::clean_names()
```

build 'region' variable from 'geography'. This is gonna be the only categorical variable in the dataset.
```{r}
data_pre = data %>% 
separate(., geography, into = c('county', 'state'), sep = ', ') %>% 
  mutate(., region = replace(state, state == 'Connecticut', 'Northeast'),
            region = replace(region, region == 'Maine', 'Northeast'),
            region = replace(region, region == 'Massachusetts', 'Northeast'),
            region = replace(region, region == 'New Hampshire', 'Northeast'),
            region = replace(region, region == 'Rhode Island', 'Northeast'),
            region = replace(region, region == 'Vermont', 'Northeast'),
            region = replace(region, region == 'New Jersey', 'Northeast'),
            region = replace(region, region == 'New York', 'Northeast'),
            region = replace(region, region == 'Pennsylvania', 'Northeast'),
            region = replace(region, region == 'Indiana', 'Midwest'),
            region = replace(region, region == 'Michigan', 'Midwest'),
            region = replace(region, region == 'Illinois', 'Midwest'),
            region = replace(region, region == 'Ohio', 'Midwest'),
            region = replace(region, region == 'Wisconsin', 'Midwest'),
            region = replace(region, region == 'Iowa', 'Midwest'),
            region = replace(region, region == 'Kansas', 'Midwest'),
            region = replace(region, region == 'Minnesota', 'Midwest'),
            region = replace(region, region == 'Missouri', 'Midwest'),
            region = replace(region, region == 'Nebraska', 'Midwest'),
            region = replace(region, region == 'North Dakota', 'Midwest'),
            region = replace(region, region == 'South Dakota', 'Midwest'),
            region = replace(region, region == 'Delaware', 'South'),
            region = replace(region, region == 'Florida', 'South'),
            region = replace(region, region == 'Georgia', 'South'),
            region = replace(region, region == 'Maryland', 'South'),
            region = replace(region, region == 'North Carolina', 'South'),
            region = replace(region, region == 'South Carolina', 'South'),
            region = replace(region, region == 'Virginia', 'South'),
            region = replace(region, region == 'District of Columbia', 'South'),
            region = replace(region, region == 'West Virginia', 'South'),
            region = replace(region, region == 'Alabama', 'South'),
            region = replace(region, region == 'Kentucky', 'South'),
            region = replace(region, region == 'Mississippi', 'South'),
            region = replace(region, region == 'Tennessee', 'South'),
            region = replace(region, region == 'Arkansas', 'South'),
            region = replace(region, region == 'Louisiana', 'South'),
            region = replace(region, region == 'Oklahoma', 'South'),
            region = replace(region, region == 'Texas', 'South'),
            region = replace(region, region == 'Arizona', 'West'),
            region = replace(region, region == 'Colorado', 'West'),
            region = replace(region, region == 'Idaho', 'West'),
            region = replace(region, region == 'Montana', 'West'),
            region = replace(region, region == 'Nevada', 'West'),
            region = replace(region, region == 'New Mexico', 'West'),
            region = replace(region, region == 'Utah', 'West'),
            region = replace(region, region == 'Wyoming', 'West'),
            region = replace(region, region == 'Alaska', 'West'),
            region = replace(region, region == 'California', 'West'),
            region = replace(region, region == 'Hawaii', 'West'),
            region = replace(region, region == 'Oregon', 'West'),
            region = replace(region, region == 'Washington', 'West'))
```


There are 3 variables that has missing data in the original dataset: 'pct_some_col18_24', 'pct_employed16_over', 'pct_private_coverage_alone'. Since 'pct_no_hs18_24', 'pct_some_col18_24', 'pct_hs18_24', and 'pct_bach_deg18_24' together equals 100, here I re-calculate the 'pct_some_col18_24' variable so that it would not contain missing data anymore.

```{r}
data_pre = 
  mutate(data_pre, pct_some_col18_24 = 100 - (pct_no_hs18_24 + pct_hs18_24 + pct_bach_deg18_24))
```

Dataset after variable pre-selection:
```{r}
data_vrb_pre_selec = data_pre %>%
  select(., -'avg_deaths_per_year', -'binned_inc', -'pct_married_households',
            -'pct_employed16_over', -'pct_private_coverage_alone', -'county', -'state')
```

```{r}
final_model = lm(formula = target_death_rate ~ incidence_rate + poverty_percent + median_age_female +
     region + pct_hs18_24 + pct_hs25_over + pct_unemployed16_over + pct_public_coverage,
     data = data_vrb_pre_selec) %>%
  summary(.)

final_mse = (final_model$sigma)^2

(final_mse)
```


CV,repeat 10 times
```{r }
mean_mse_cv = 0

for(i in 1:10){
  data_train<-trainControl(method="cv", number=10)

  model_caret<-train(target_death_rate ~ incidence_rate + poverty_percent + median_age_female +
     region + pct_hs18_24 + pct_hs25_over + pct_unemployed16_over + pct_public_coverage, 
                   data = data_vrb_pre_selec,
                   trControl=data_train,
                   method='lm',
                   na.action=na.pass)
  
  MSE = mean((model_caret$resample$RMSE)^2)

  mean_mse_cv = mean_mse_cv + MSE/10
}

(mean_mse_cv)
```

Test MSE is similar to the MSE of our final model, so the predictive ability of our model is good.